### 2022-04-21 manuscript.tex
fig name update
Updates from Overleaf

Conferences are spaces to meet and to network within and across academic and technical fields, to learn about new advances, and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces ten simple rules to host an inclusive conference based on the authors' recent experience organizing the 2021 edition of the useR! statistical computing conference, which attracted a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been written mainly for potential organizers and selection committees of conferences and contain multiple practical tips to help a variety of events become more accessible and inclusive. We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
\hl{* Translated versions of the extended abstract are available in ten languages in the Supplementary Information: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.}

@@

% Roadmap of the paper
Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference.
\linelabel{methods-a} \hl{The rules were first drafted by the core team and the members of the diversity, accessibility, and inclusion team of useR! 2021 who wrote their own ten simple rules (see Supplementary Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
A final list was proposed and then discussed and reviewed by the rest of the coauthors.} \linelabel{methods-b}
The rules are organized in three sections: three foundation rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
The \textbf{foundation rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
The \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and thematic sessions). 
\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
In \textbf{Rule 7}, we provide suggestions to account for linguistic diversity. 
\textbf{Rule 8} offers tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion and for passing the torch to future organizers.

@@

\begin{figure}[!h]
\centering
%ast: use Adobe Acrobat to add missing accessibility metadata to your PDF file. ask how to do this this when submitting - some edits to the PDF may be needed to add alt-text to the final version https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#late
\pdftooltip{
\includegraphics[width=\textwidth]{figs/fig1_rules.png}}{
%Alt-text for the figure to be included in the final PDF: Diagram of how the 10 rules are organized into three groups: foundation rules (Rules 1-3), design rules (Rules 4 to 9), and a continuity rule (Rule 10). The overall diagram depicts an arrow from left to right which contains all the rules. In the left column, under the first label" Foundation", there are three text boxes in light blue: "1. Vision", "2. Safe environment", and "3. Organizing Team". In the middle column, under the label "Design Rules" there are six light green boxes with Rules "4. Counter bias", "5. Virtual / hybrid", "6. Accessibility", "7. Language", "8. Communication",  "9. Finance". To the right, below the arrowhead labeled Continuity, is the Tenth Rule in a light pink box "10. Pass the torch".
}
\caption{Schematic diagram of the rules organized in three groups: foundation (Rules 1, 2, and 3), design (Rules 4 to 9), and continuity (Rule 10).}
\label{fig:diagram}
\end{figure}

@@

\begin{figure}[!h]
\centering
\pdftooltip{
\includegraphics[width=\textwidth]{figs/marmots.pdf}}{
}
\caption{Margot the marmot was the useR! 2021 mascot. To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing the colors of diverse communities of practice that are part of the large R community and specific groups of people we wanted to support explicitly. From left to right: Margot wearing scarves with the colors of MiR (Minorities in R \url{https://mircommunity.com}), R-Ladies \url{https://rladies.org}, AfricaR \url{https://africa-r.org}, and the LGBTQIA+ and transgender communities. Margot was created by Francisco Etchart and is available under a CC-BY-NC-SA license (\url{https://github.com/useRconf/visuals-2021}).}
\label{fig:marmots}
\end{figure}
%alt-text for the figure, to be included in the final PDF: Five marmots with colorful scarves. From left to right: a marmot with a dark and light green scarf from the MiR community is watching her tablet, a marmot with the purple and gray scarf for R-Ladies is stepping on a rock and watches the horizon, a marmot with an orange and purple AfricaR scarf is sitting on a couch watching a film on her laptop with popcorn, a yogi marmot with a rainbow LGBTQIA+ scarf does the warrior pose on her blue mat, and finally, a marmot with a scarf with the pink, light blue and white color of the transgender pride flag waves her hand.

@@

To list author contributions, we used the high-level contributor roles in the CRediT framework that we considered relevant for this study (project administration, supervision, writing original draft, conceptualization, writing—major reviews and edits, resources, and visualization; \url{http://credit.niso.org}), and additional categories to acknowledge those who shared their own ten simple rules for an inclusive conference without looking at the first draft (i.e. sharing rules), and those who translated the extended abstract into different languages.

@@

\section*{Acknowledgments}
The authors of this piece would like to thank every single member of the organizing team of useR! 2021 (\url{https://user2021.r-project.org/about/global-team}) for their valuable contribution to an inclusive conference experience and the R Foundation for trusting us with the organization of useR! 2021 and supporting us through the process. Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, and Francisco Etchart for his help with the figures. We would also like to thank Arjun Krishnan, Julia Ganz, Ingo Braasch, Susan Ewart, and Hilda Mejia-Abreu for their feedback on the manuscript. In addition to some of the authors, we would like to acknowledge the following contributors for translating our extended abstract to several languages: Koki Tsuyuzaki and Kozo Nishida [Japanese], KwangChun Lee [Korean], Arjun Krishnan [Tamil], Kewalin Samart, Natnaree Yolnava, and Wanprakai Kaewphrae [Thai].
The comments of the anonymous reviewers contributed to the improvement of the manuscript and we thank them for that.

@@

HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article and by The R Foundation for Statistical Computing for the organization of useR! 2021. HT is a board member of The R Foundation for Statistical Computing, serving on the R Foundation conference committee that arranges hosts for the useR! conference series and chairing Forwards, the R Foundation taskforce for women and under-represented groups. The R Foundation is currently funding YBS to support the useR! 2022 organization team. HT was administrator and volunteer for the R Project Google Season of Docs 2021 project, that funded work by AST (writer) and YBS (steering committee).
The organization of useR! 2021 was largely a volunteer effort, but stipends were available from the conference funds for specific roles. RJ, AST, SM, BA, NM, AU, and AE received code of conduct team stipends; AST, YBS, JPNG, MAC, RG, and JC received zoom host stipends; AST and MAC received stipends for their unrelenting efforts as diversity and social program leads, respectively. JR is funded by start-up award from Michigan State University; publication cost will waived due to PLoS-MSU partnership.

\section*{Data availability}
For the purpose of open access, the authors have applied a Creative Commons Attribution (CC BY) license to any Author Accepted Manuscript version arising from this submission. The manuscript and the supporting docs are available at: https://github.com/rociojoo/Ten-simple-rules-inclusive-conference.

@@

>>>>

Conferences are spaces to meet and to network within and across academic and technical fields, to learn about new advances, and share our work. They can help define career paths and create long-lasting collaborations and opportunities. 
However, these opportunities are not equal for all. 
This article introduces ten simple rules to host an inclusive conference based on the authors' recent experience organizing the 2021 edition of the useR! statistical computing conference, which attracted a broad range of participants from academia, industry, government, and the non-profit sector. 
Coming from different backgrounds, career stages, and even continents, we embraced the challenge of organizing a high-quality virtual conference in the context of the COVID-19 pandemic and making it a kind, inclusive, and accessible experience for as many people as possible.
The rules result from our lessons learned before, during, and after the organization of the conference. 
They have been written mainly for potential organizers and selection committees of conferences and contain multiple practical tips to help a variety of events become more accessible and inclusive. We see this as a starting point for conversations and efforts towards building more inclusive conferences across the world.
\hl{* Translated versions of the English abstract and the list of rules are available in ten languages in the Supporting Information: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.}

@@

% Roadmap of the paper
Here, we present a set of ten rules based on the lessons we learned before, during, and after the organization of the useR! conference.
\linelabel{methods-a} \hl{The rules were first drafted by the core team and the members of the diversity, accessibility, and inclusion team of useR! 2021 who wrote their own ten simple rules (see Supporting Information) based on the collective experience of organizing the conference and fueled by personal experiences and literature review.
A final list was proposed and then discussed and reviewed by the rest of the coauthors.} \linelabel{methods-b}
The rules are organized in three sections: three foundation rules, six design rules, and a continuity rule (\textbf{Fig. \ref{fig:diagram}}).
The \textbf{foundation rules} comprise key elements to conceive the work on diversity and inclusion in any conference. 
\textbf{Rule 1} is about setting a vision of diversity and inclusion that should guide all the efforts and decision making in the organization.
\textbf{Rule 2} focuses on how to create a safe and welcoming environment for all the attendees. 
\textbf{Rule 3} highlights the importance of starting with an inclusive and diverse organizing team and provides tips on work dynamics.
The \textbf{design rules} focus on weaving inclusion into the conference design process.
In \textbf{Rule 4}, we introduce multiple ways to counteract bias in the conference program (keynotes, program committee, abstract selection, and thematic sessions). 
\textbf{Rule 5} provides advice for designing an inclusive online component in virtual and hybrid conferences.
\textbf{Rule 6} focuses on accessibility practices to include people with disabilities.
In \textbf{Rule 7}, we provide suggestions to account for linguistic diversity. 
\textbf{Rule 8} offers tips for developing an inclusive communication strategy. 
In \textbf{Rule 9}, we address budgeting for inclusive practices and helping participants with affordable registration costs, scholarships, and other forms of financial support.
Finally, 
\textbf{Rule 10}, the \textbf{continuity} rule, emphasizes the importance of self-assessment and advocates for making the conference part of a long-term commitment to inclusion and for passing the torch to future organizers.

@@

\begin{figure}[!h]
\centering
%ast: use Adobe Acrobat to add missing accessibility metadata to your PDF file. ask how to do this this when submitting - some edits to the PDF may be needed to add alt-text to the final version https://chi2021.acm.org/for-authors/presenting/papers/guide-to-an-accessible-submission#late
\pdftooltip{
\includegraphics[width=\textwidth]{figs/fig1.png}}{
%Alt-text for the figure to be included in the final PDF: Diagram of how the 10 rules are organized into three groups: foundation rules (Rules 1-3), design rules (Rules 4 to 9), and a continuity rule (Rule 10). The overall diagram depicts an arrow from left to right which contains all the rules. In the left column, under the first label" Foundation", there are three text boxes in light blue: "1. Vision", "2. Safe environment", and "3. Organizing Team". In the middle column, under the label "Design Rules" there are six light green boxes with Rules "4. Counter bias", "5. Virtual / hybrid", "6. Accessibility", "7. Language", "8. Communication",  "9. Finance". To the right, below the arrowhead labeled Continuity, is the Tenth Rule in a light pink box "10. Pass the torch".
}
\caption{Schematic diagram of the rules organized in three groups: foundation (Rules 1, 2, and 3), design (Rules 4 to 9), and continuity (Rule 10).}
\label{fig:diagram}
\end{figure}

@@

\begin{figure}[!h]
\centering
\pdftooltip{
\includegraphics[width=\textwidth]{figs/fig2.pdf}}{
}
\caption{Margot the marmot was the useR! 2021 mascot. To represent the global and inclusive nature of the conference, Margot was drawn using scarves representing the colors of diverse communities of practice that are part of the large R community and specific groups of people we wanted to support explicitly. From left to right: Margot wearing scarves with the colors of MiR (Minorities in R \url{https://mircommunity.com}), R-Ladies \url{https://rladies.org}, AfricaR \url{https://africa-r.org}, and the LGBTQIA+ and transgender communities. Margot was created by Francisco Etchart and is available under a CC-BY-NC-SA license (\url{https://github.com/useRconf/visuals-2021}).}
\label{fig:marmots}
\end{figure}
%alt-text for the figure, to be included in the final PDF: Five marmots with colorful scarves. From left to right: a marmot with a dark and light green scarf from the MiR community is watching her tablet, a marmot with the purple and gray scarf for R-Ladies is stepping on a rock and watches the horizon, a marmot with an orange and purple AfricaR scarf is sitting on a couch watching a film on her laptop with popcorn, a yogi marmot with a rainbow LGBTQIA+ scarf does the warrior pose on her blue mat, and finally, a marmot with a scarf with the pink, light blue and white color of the transgender pride flag waves her hand.

@@

To list author contributions, we used the high-level contributor roles in the CRediT framework that we considered relevant for this study (project administration, supervision, writing original draft, conceptualization, writing—major reviews and edits, resources, and visualization; \url{http://credit.niso.org}), and additional categories to acknowledge those who shared their own ten simple rules for an inclusive conference without looking at the first draft (i.e. sharing rules), and those who translated the English abstract and the list of rules into different languages.

@@

\section*{Acknowledgments}
The authors of this piece would like to thank every single member of the organizing team of useR! 2021 (\url{https://user2021.r-project.org/about/global-team}) for their valuable contribution to an inclusive conference experience and the R Foundation for trusting us with the organization of useR! 2021 and supporting us through the process. Special thanks to Saranjeet Kaur and gwynn sturdevant for suggestions to previous versions of the manuscript, and Francisco Etchart for his help with the Marmot figures. We would also like to thank Arjun Krishnan, Julia Ganz, Ingo Braasch, Susan Ewart, and Hilda Mejia-Abreu for their feedback on the manuscript. In addition to some of the authors (listed in the Translated Abstracts section of Supporting Information), we would like to acknowledge the following contributors for translating our extended abstract: Koki Tsuyuzaki and Kozo Nishida [Japanese], KwangChun Lee [Korean], Arjun Krishnan [Tamil], Kewalin Samart, Natnaree Yolnava, and Wanprakai Kaewphrae [Thai].
The comments of the anonymous reviewers contributed to the improvement of the manuscript and we thank them for that.

@@

JR is funded by start-up award from Michigan State University (MSU); publication costs were waived because of MSU’s institutional partnership with PLoS. HT was supported by the Engineering and Physical Sciences Research Council [EPSRC EP/V052128/1] during the preparation of this article.

The organization of useR! 2021---which inspired this work---was largely a volunteer effort (from many people including all the authors), but stipends were given from the conference funds for some specific roles.
AE, AU, AST, BA, JC, JPNG, MAC, NM, RG, RJ, SM, and YBS received stipends for roles that include code of conduct team, zoom host, and diversity and social program leads.

\section*{Data availability}
For the purpose of open access, the authors have applied a Creative Commons Attribution (CC BY) license to any author accepted manuscript version arising from this submission. The manuscript, figures, supplementary material, and supporting docs are available at: https://github.com/rociojoo/Ten-simple-rules-inclusive-conference.

@@

\section*{Supporting Information}

Supporting Information 1. Abstract and list of rules in multiple languages: Arabic, French, German, Italian, Japanese, Korean, Portuguese, Spanish, Tamil, and Thai.

Supporting Information 2. Rules proposed in the process of shaping these ten simple rules.
